\documentclass[../main]{subfiles}

\begin{document}

\section{Halogenoalkanes}

	\subsection{Structure of Halogenoalkanes}

	Halogenoalkanes are derivatives of alkanes compounds with \ch{F, Cl, Br, I} groups. Halogenarenes are derivatives of arenes with benzylic \ch{H} replaced by a \ch{F, Cl, Br, I} group. \\

	Halogenoalkanes can be classified according to the degree of substitution of \ch{C} that the halogen is bonded to.

	\subsection{Physical Properties of Halogenoalkanes}

	Halogenoalkanes contain polar \ch{R-X} bonds which increase their melting and boiling point. For compounds with the same \ch{R} group, the molecule with a larger \ch{X} electron cloud has stronger id-id interactions and hence a larger melting and boiling point. \\

	Though they are polar, halogenoalkanes are still poorly soluble in water. Fluoroalkanes and monochloroalkanes are less dense than water while other \ch{R-X} compounds are denser than water.

	\subsection{Formation of Halogenoalkanes}

	\noindent \textbf{Alkanes to Halogenoalkanes}

	Mono and polyhalogenated products can be formed from alkanes. Refer to `Free Radical Substitution' in Alkanes. \\

	\noindent \textbf{Alkenes to Halogenoalkanes}

	Mono and dihalogenated products can be formed from alkenes. Refer to `Electrophilic Substitution' in Alkenes. \\

	\noindent \textbf{Alcohols to Halogenoalkanes}

	Refer to `Nucleophilic Substitution' in Hydroxy Compounds. \\

	\subsection{Reaction of Halogenoalkanes}

	Halogenoalkanes typically undergo nucleophilic substitution reactions.

	The rate of reaction primarily depends on the nature of the \ch{R-X} bond, with weaker bonds reacting faster despite the change in electronegativity, hence rate of reaction increases in the order \ch{C-Cl},\ch{C-Br} and \ch{C-I}. \\

	\subsubsection{Nucleophilic Substitution}

	\scidef{Nucleophile}{A Nucleophile \ch{Nu} is a particle which has at least one pair of free electrons and can act as an electron donor. Nucleophiles are attracted to electron-poor sites.}

	Note that nucleophiles are not necessarily charged. \\

	Nucleophilic substitution can occur through two mechanisms, characterized by the study of these mechanism's kinetic properties.

	\scidef{\usub{S}{N}2}{The \usub{S}{N}2 mechanism of nucleophilic substitution involves two molecules in its slow (and only) step.}

	\usub{S}{N}2 reactions involve the backside attack of a nucleophile to the halogenoalkane at the opposite side of the \ch{C-X} bond, forming an intermediate molecule both bonded to the nucleophile and \ch{X}, after which \ch{X} is ejected. This reaction preserves the enantiomeric quality of a molecule if the original alkane displays enantiomerism about the \ch{C} with \ch{X} due to the spatial requirements of a backside attack. 

	\chfg{
		\subscheme[-90]{
			\subscheme{
				\chemfig{@{a1}\lewis{0:,\ch{HO-}}} \arrow{0}[,0] \+ \chemfig*{@{a2}C(-[@{a3}:0]@{a4}Br)(-[:115]H)(<:[:205]H3C)(<[:250,,,1]CH2CH3)} 
				\chemmove[shorten <=2pt,shorten >=2pt]{\draw (a1.north east)..controls +(45:6mm) and +(150:4mm) ..(a2); }
				\chemmove[shorten <=2pt,shorten >=2pt]{\draw (a3)..controls +(90:5mm) and +(90:5mm) ..(a4); }
			}
			\arrow{0}[,0.2]
			\subscheme{
				\arrow \chemleft[ \chemfig*{\chemabove{HO}{\delta-}-[,,,,dash bond]C(-[:90]H)(<:[:250]H3C)(<[:290]CH2CH3)-[,,,,dash bond]\chemabove{HO}{\delta-}} \chemright{]^{\text{\textdoublebarpipe}}}
			}
			\arrow{0}[,0.2]
			\subscheme{\arrow \chemfig*{HO-C(-[:65]H)(<:[:335]CH3)(<[:290]CH2CH3)} \arrow{0}[,0] \+ \ch{Br-}}
		}
	} \\

	Label partial charges on atoms as they `develop' charges throughout the reaction. A negatively charged nucleophile will progress to a neutral charge whereas a neutrally charged nucleophile will develop a positive charge.

	\scidef{\usub{S}{N}1}{The \usub{S}{N}1 mechanism of nucleophilic substitution involves 1 molecule in its slow step.}

	\usub{S}{N}1 reactions involve the spontaneous ejection of the \ch{X} group in its slow step to form a carbocation which will then react with a nucleophile. The later addition of the nucleophile will happen at equal probabilities at either side of the planar \ch{C+}, hence any enantiomeric products about \ch{C} will be racemic. \\

	\chfg{
		\chemfig*[][scale=0.8]{C(-[@{a1}:0]@{a2}Br)(-[:115]H)(<:[:205]H3C)(<[:250,,,1]CH2CH3)} \arrow{->[slow]} \chemfig*[][scale=0.8]{\ch{C+}(-[:90]H)(<:[:250]H3C)(<[:290,,,1]CH2CH3)} \arrow{0}[,0] \+ \ch{Br-}
		\chemmove[shorten <=2pt,shorten >=2pt]{\draw (a1)..controls +(90:5mm) and +(90:5mm) ..(a2); }
	} 

	\chfg{
		\chemfig{@{a1}\lewis{0:,\ch{HO-}}} \arrow{0}[,0] \+ \chemfig*[][scale=0.8]{@{a2}\ch{C+}(-[:90]H)(<:[:250]H3C)(<[:290,,,1]CH2CH3)} \arrow{0}[,0] \+ \chemfig{@{a3}\lewis{4:,\ch{OH-}}}
	}

	\begin{tikzpicture}[node distance=0cm and 2cm]
		\node (a1) {};
		\node[above right=of a1] (a2) {
			\chemname{\chemfig*[][scale=0.8]{HO-C(-[:65]H)(<:[:335]CH3)(<[:290]CH2CH3)}}{Inversion of Configuration}
			};
		\node[below right=of a1] (a3) {
			\chemname{\chemfig*[][scale=0.8]{C(-[:115]H)(<:[:205]CH3)(<[:250,,,1]CH2CH3)-OH}}{Retention of Configuration}
			};
		\draw[-stealth] (a1) -- ( $ (a1.0)!0.5!(a2.west|-a1.0) $ ) |- (a2.west) node[auto,pos=0.7] {};
		\draw[-stealth] (a1) -- ( $ (a1.0)!0.5!(a3.west|-a1.0) $ ) |- (a3.west) node[auto,pos=0.7] {};
	\end{tikzpicture}

	Draw stereochemistry of a reaction if the original molecule displays enantiomerism about the \ch{C} bonded to \ch{X}. Also indicate by annotating `Retention of Configuration' and `Inversion of Configuration' below the resultant molecules. \\

	The mechanism at which a \usub{S}{N} reaction occurs depends primarily on the bonding of the \ch{C} bonded to \ch{X}. \ch{C} with more alkyl groups bonded to it will generally follow the \usub{S}{N}{1} mechanism because it will form a more stable \ch{C+} whereas \ch{C} with less alkyl groups bonded to it will follow the \usub{S}{N}2 mechanism due to the free space around \ch{C} leaving it susceptible to backside attack. When explaining why a reaction follows one mechanism, explain why one is favored and why the other is less likely. \\

	\noindent \textbf{Formation of Alcohol}

	Reagents: \ch{NaOH} \\
	Conditions: Aqueous, heat \\

	A polyhalogenoalkane may react to form multiple alcohol groups on the same carbon (also known as gem-diols), which may then decompose into ketone and carboxylic acid groups. \\

	\noindent \textbf{Formation of Amines}

	Reagents: Excess concentrated \ch{NH3} \\
	Conditions: Alcoholic, heat, sealed tube \\

	If \ch{NH3} is not in excess, a primary amine can act as a nucleophile in later reactions and form a chain reaction to eventually form secondary and tertiary amines and finally a quaternary ammonium salt because more-substituted amine compounds are stronger nucleophiles than their primary counterparts. \\

	Indicate all reactions (including proton transfer) when describing the mechanism. \\

	\noindent \textbf{Formation of Nitriles}

	Reagents: \ch{KCN} \\
	Conditions: Alcoholic, heat \\

	In practice, a mixture of alcohol and water is used so that all reactants can be dissolved. Nitriles can be further reduced to form primary amines (\ch{H2}, \ch{Ni} and Heat OR \ch{LiAlH4-}, dry ether and heat) or undergo hydrolysis to form carboxylic acids (\ch{H2SO4 (aq)}, heat) and carboxylate ions (\ch{NaOH (aq)}, heat). \\

	\noindent \textbf{Williamson Ether Synthesis}

	Reagents: Alcohol, \ch{Na} OR \ch{R-O- Na+}\\
	Conditions: Heat \\

	\subsubsection{Elimination}

	Though the benzene ring itself is unreactive, any alkyl chains bonded to a benzene ring can undergo oxidation, replacing the alkyl group with a \ch{COOH} group. However, the reaction requires that the benzyllic carbon has at least 1 \ch{H} or 1 \ch{O} atom bonded to it, otherwise a case such as a tert-butyl benzene will not be oxidised due to steric effect of \ch{CH3} groups attached to the benzylic \ch{C}. \\

	\noindent \textbf{Dehydrohalogenation}

	Reagents: \ch{NaOH} \\
	Conditions: Alcoholic, heat \\

	\chfg{
		\subscheme[-90]{
			\subscheme{
				\chemfig{-[:-30](-[@{a1}:270]@{a2}Br)-[@{a3}:30](-[@{a4}:90]@{a5}H)-[:-30]} \+ \chemfig{@{a6}\lewis{2:,O}\ch{H-}}
				\chemmove[shorten <=2pt,shorten >=2pt]{ \draw (a1)..controls +(180:5mm) and +(180:5mm) ..(a2); }
				\chemmove[shorten <=2pt,shorten >=2pt]{ \draw (a4)..controls +(180:5mm) and +(120:5mm) ..(a3); }
				\chemmove[shorten <=2pt,shorten >=2pt]{ \draw (a6)..controls +(90:5mm) and +(0:5mm) ..(a5); }
			}
			\arrow{0}[,0.2]
			\subscheme{\arrow \chemfig{-[:-30]=[:30]-[:-30]} + \ch{H2O} \+ \ch{Br-}}
		}
	}

	Dehydrohalogenation can only occur if there is a \ch{H} in some \ch{C} adjacent to the \ch{C-X} group. \\

	\subsection{Halogenoarenes}

	Halogenoarenes are compounds with \ch{X} bonded to benzene rings. \\

	Halogenoarenes burn with a sooty flame and are insoluble in and denser than water. \\

	Refer to `Electrophilic Substitution' in Arenes for methods of preparation. \\

	Halogenoarenes generally do not undergo nucleophilic substitution reactions. Extension of the \(\pi\) \ch{e-} cloud over \ch{C-X} causes the formation of a partial double bond character, the benzene ring physically blocks approaching nucleophiles, the \ch{e-} density of the benzene ring repels approaching nucleophiles and the presence of the \ch{e-} cloud prevents the formation of \ch{C+}. However, reactions may occur under extreme conditions (\SI{150}{\atm}, \SI{350}{\celsius}).\\

	Halogenoarenes can also undergo electrophilic substitution, much like other arenes. However, \ch{X} are electron-withdrawing and electronegative, hence they are deactivating groups and the reactivity of the arene decreases as more \ch{X} groups are added. \ch{X} is a 2,4-directing group.

	\subsection{Distinguishing Tests}

	\subsubsection{`4-step Test'}

	\begin{enumerate}
		\item Heat with \ch{NaOH (aq)} (liberate \ch{-X} group)
		\item Cool (prevent decomposition of \ch{AgNO3})
		\item Acidify with \ch{HNO3} (remove OH)
		\item Add \ch{AgNO3} (form ppt)
	\end{enumerate}

	Positive tests will form a \ch{AgX} precipitate. The identity of \ch{X} can be determined according to the color of the precipitate. \\

	\subsubsection{Ethanolic \ch{AgNO3}}

	\begin{enumerate}
		\item Heat with ethanolic \ch{AgNO3}
	\end{enumerate}

	Ethanol acts as a nucleophile which will displace the halide group, which then causes the precipitation of \ch{AgX}.

	\subsection{Applications of Halogenoalkanes}

	Strong \ch{C-F} and \ch{C-Cl} bonds mean that organic molecules saturated with \ch{F} and \ch{Cl} are highly unreactive. As a result, Chlorofluorocarbon (CFC) / `Freon' compounds are also usually non-toxic, non-flammable (or even flame retardant), odorless, volatile and easily liquifiable. However, their high stability and low density mean that these compounds can rise to high altitudes before decomposition, after which \ch{Cl^.} radicals can easily form to deplete \ch{O3} in a chain reaction. \\

	As a replacement, the use of HCFC compounds and HFC compounds are proposed as the presence of \ch{C-H} bonds allow them to be destroyed at lower altitudes before being able to extensively damage the \ch{O3} layer.
	
\end{document}