\documentclass[../../main]{subfiles}

\begin{document}

\subsection{Purpose and Value}

Tl;dr: To ``Inform, ediucate and entertain'' \\

Purpose and value are two very distinct sets of questions, but still govern the questions asking the whys of media. Purpose itself pertains to the idealistic use case of different forms of media and less so of how efficient they are at achieving goals, whereas value pertains to the worth of the existence and use of a form of media be it originating from the way it is conveyed to the content that is conveyed to the effectiveness of the form of media, typically alternating between different timeframes of past success and future potential of a form of media to argue for one's value or lack thereof. 

\subsubsection{Purpose}

\paragraph{Consider the argument that the main purpose of television should be to educate than simply to entertain. (GCE-A/15)}-

\paragraph{How far is the media responsible for promoting democracy in your society? (RJC/14)}-

\paragraph{DYA that the mass media should pursue responsibility and not profit? (RJC/11)}-

\subsubsection{Value}

\paragraph{Consider the view that spoken language is more important than the written form. (GCE-A/13)         }-

\paragraph{Do films offer anything more than an escape from reality? (GCE-A/14)} -

\paragraph{Is there any value in horror films and books? (RJC/14)}-

\paragraph{Are bookstores still relevant in today's world? (RJC/13)}-

\paragraph{TWE does the media create mediocrity? (RJC/11)}-

\end{document}