\documentclass[../main]{subfiles}

\begin{document}

\section{Oscillations}

\subsection{Simple Harmonic Motion}

\scidef{Simple Harmonic Motion}{Simple Harmonic Motion (SHM) occurs when a body oscillates about a point where its acceleration is proportional to its displacement from said point and directed towards said point.}

\subsection{SHM Equations}

\scieqn[gathered]{SHM Equation}{For }{
	a = -kx \\
	x = x_o \cos{\omega t} \\
	v = \frac{dx}{dt} = - x_o \omega \sin{\omega t} \\
	a = \frac{dv}{dt} = - x_o \omega^2 \cos{\omega t} \\
	  = - \omega^2 x \\
	k = \omega^2
}

\subsubsection{Graphs of SHM}

Bodies undergoing SHM have sinusoidal graphs of displacement, velocity and acceleration due to the trigonometric functions in their equations. Graphs of energy of SHM hold the property of conservation of energy, where the sum of potential energy and kinetic energy, usually in the form of square trigonometric function graphs, always add up to a constant.

\subsection{Cases of SHM}

\subsubsection{Horizontal Oscillations}

Horizontal Spring Systems involve a mass connected to a spring. The spring creates a horizontal restoring force.

Pendulum Systems involves a mass connected to a inextensible string. The net force of gravity and tension in the string creates a horizontal restoring force which is assumed to be proportional to horizontal displacement at small angles \(\theta < 6 \degree \). 

\subsubsection{Vertical Oscillations}

Vertical Spring Systems involve a mass connected to a vertical spring. The net force of the spring and gravity provides the restoring force which acts on the oscillating mass. The center of oscillations is assumed to be the same position as when the mass is in stationary equilibrium and any SHM acts around this position. Energy is also conserved, where the summation of spring potential and gravitational potential stores energy to be converted into kinetic energy.

\subsection{Damping}

\scidef{Damping}{Damping occurs when there is an external force acting on the object undergoing SHM, usually proportional to velocity (such as air resistance).}

\scidef{Light Damping}{Light Damping occurs when a damping force causes the amplitude of SHM to decrease exponentially over time.}

\scidef{Critical Damping}{Critical Damping occurs when a damping force causes a body undergoing SHM to return to equilibrium position and stop oscillating within the shortest possible time.}

\scidef{Heavy Damping}{Heavy Damping occurs when a damping force causes a body undergoing SHM to return to equilibrium position and stop oscillating over a period of time longer than if the system were critically damped.}

\subsubsection{Forced Oscillations}

\scidef{Forced Oscillations}{Forced Oscillations arise when a system has to receive external force in order to undergo SHM.}

\subsection{Resonance}

\scidef{Resonance}{Resonance occurs when a system responds to a driving force with a maximum amplitude. This implies the maximal transfer of energy between driving and driven systems, therefore implying that the driving frequency is equal to the natural frequency of the driven system.}

\scidef{Natural Frequency}{The Natural Frequency, also known as the Resonant Frequency or Resonance Frequency, is the frequency at which a system is able to receive energy at a maximum.}

RI's tutorial solutions have used the terms ``Resonance Frequency'', ``System at Resonance'' and ``Resonance Occurs''. Use these if phrasing is vague and if you worry about presentation / phrasing marks.

\end{document}