\documentclass[../../main]{subfiles}

\begin{document}

\subsection{Education}

\subsubsection{Medium of Education}

\paragraph{`Books serve little purpose in education as technological developments become more sophisticated.' How far DYA? (GCE-A/15)} See section 1.2 New Media

\paragraph{Consider the argument that the main purpose of television should be to educate than simply to entertain. (GCE-A/15)} See section 1.1 Purpose of Media


\subsubsection{Manner of Education}

\paragraph{`Only grades matter in education.' Discuss. (RJC/14)}-

\paragraph{`Intelligence plus character  - that is the goal of true education.' Discuss. (RJC/13)}-

\paragraph{TWE should schools use examinations to evaluate students? (RJC/12)}-


\subsubsection{Matter of Education}

\paragraph{`Education should only be concerned with what is useful in life.' Discuss. (GCE-A/13)}-

\paragraph{Is the study of history increasingly losing its relevance in today's world? (RJC/15)}-

\paragraph{How important is the study of History for a nation's future? (RJC/15)}-

\paragraph{Consider the view that the study of history is essential for mankind. (RJC/14)}-

\paragraph{TWE do you think that creativity can be cultivated? (RJC/14)}-

\paragraph{How relevant is the study of History in our world today? (RJC/14)}-

\paragraph{`The teaching of literary classics should be made compulsory in schools.' Discuss. (RJC/13)}-

\paragraph{`Singapore can do without the study of the humanities.' Discuss. (RJC/13)}-


\subsubsection{Impacts of Education}

\paragraph{Does better education necessarily lead to a better standard of living? (GCE-A/11)}-

\paragraph{`Education has produced a population that is literate but not compassionate.' How far is this true of your country? (RJC/15)} See section 2.3 Social Stratification

\paragraph{`Qualified but not enlightened.' Is this a fair description of educated people today? (RJC/13)}-

\paragraph{`Education has resulted in only more inequality.' DYA? (RJC/11)}

\end{document}